---
title: "Statistical Analysis"
output:
  github_document:
    code_folding: hide
---

<style>
div.blue { background-color:aliceblue; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

##### On this page, we explored how walkability is correlated with health outcomes at the national level.

</div>

<br>

### Setup

To ease interpretation, the walkability score was dichotomized according to [cutoff values](https://www.epa.gov/sites/default/files/2021-06/documents/national_walkability_index_methodology_and_user_guide_june2021.pdf) provided by EPA, where values above 10.5 are classified as "More Walkable", and those below as "Less Walkable".

<details>
  <summary>Click to show code</summary>
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(modelr)
library(betareg)
library(mgcv)
library(purrr)
library(broom)
library(mfx)
```

```{r pivot-dataset, cache=TRUE}
dataset = read_csv("./data/merge_final.csv")

regress_df = dataset %>% 
  pivot_wider(names_from = "race",
               values_from = "percent_race") %>% 
  pivot_wider(names_from = "measure_id",
              values_from = "data_value") %>% 
  janitor::clean_names() %>% 
  dplyr::select(tractce, nat_walk_ind, csa_name, lat, long, sex_male, sex_female, age_18plus, median_income, white, black, aian, asian, x2_plus, hispanic, highchol, mhlth, lpa, phlth, bphigh) %>% 
  mutate(
    sex_male = as.numeric(sex_male)/100,
    age_18plus = as.numeric(age_18plus)/100,
    white = as.numeric(white)/100,
    black = as.numeric(black)/100,
    aian = as.numeric(aian)/100,
    asian = as.numeric(asian)/100, 
    x2_plus = as.numeric(x2_plus)/100, 
    hispanic = as.numeric(hispanic)/100, 
    highchol = highchol/100, 
    mhlth = mhlth/100, 
    lpa = lpa/100, 
    phlth = phlth/100, 
    bphigh = bphigh/100,
    median_income = as.numeric(median_income)
    ) %>%
  group_by(tractce) %>% 
  summarize(
    nat_walk_tract = nat_walk_ind,
    sex_male = sex_male,
    age_18plus = age_18plus,
    white = white,
    black = black,
    aian = aian,
    asian = asian, 
    x2_plus = x2_plus, 
    hispanic = hispanic, 
    highchol = highchol, 
    mhlth = mhlth, 
    lpa = lpa, 
    phlth = phlth, 
    bphigh = bphigh,
    median_income = median_income,
    lat = lat,
    long = long
    ) %>% 
  dplyr::select(!c("lat", "long")) %>% 
  mutate(walkable = ifelse(nat_walk_tract<10.5, "Less Walkable", "More Walkable")) %>% 
  unique()
```

```{r beta-test-code}
# mod <- gam(highchol ~ nat_walk_tract + sex_male + age_18plus + black + median_income + s(lat, long, bs="gp", m=2),
#            family = betar(link='logit'),
#            data = regress_df)
# 
# summary(mod)
# 
# residual1_plot =
#     regress_df %>%
#     add_predictions(mod) %>%
#     add_residuals(mod) %>%
#     ggplot(aes(x = nat_walk_tract, y = resid)) +
#     geom_point(alpha = 0.3, size = 0.3) +
#     geom_smooth(se = FALSE, method = "lm")
# 
# residual1_plot

# mod_beta = betareg(mhlth ~ nat_walk_tract + sex_male + age_18plus + black, data = test)
# 
# summary(mod_beta)
# 
# residual_plot =
#     test %>%
#     add_predictions(mod_beta) %>%
#     add_residuals(mod_beta) %>%
#     ggplot(aes(x = nat_walk_tract, y = resid)) +
#     geom_point(alpha = 0.3) +
#     geom_smooth(se = FALSE, method = "lm")
# 
# residual_plot
```
</details>
<br>

### Regression Model {.tabset}

We utilized beta regression to model the relationship between health outcomes with the dichotomized walkability variable, controlling for sociodemographic variables of prevalence of age 18+, males, Black folks, and median income. 

We chose this regression method as the outcome variable is in proportions, and thus takes the form of a continuous value bounded between 0 and 1. The R package `betareg` was used to implement this method, while the `betamfx` command under the `mfx` package was used to obtain the changes in marginal probabilities on the outcome resulting from changes in the predictor. We will report both numbers, however, we will be interpreting the marginal probabilities for its more direct relevance to the real world.
  
``` {r regress, cache=TRUE}
# Beta regression model
chol = betareg(highchol ~ walkable + sex_male + age_18plus + black + median_income, data = regress_df) %>% broom::tidy(conf.int=TRUE)
mental = betareg(mhlth ~ walkable + sex_male + age_18plus + black + median_income, data = regress_df) %>% broom::tidy(conf.int=TRUE)
lpa = betareg(lpa ~ walkable + sex_male + age_18plus + black + median_income, data = regress_df) %>% broom::tidy(conf.int=TRUE)
physical = betareg(phlth ~ walkable + sex_male + age_18plus + black + median_income, data = regress_df) %>% broom::tidy(conf.int=TRUE)
bloodpres = betareg(bphigh ~ walkable + sex_male + age_18plus + black + median_income, data = regress_df) %>% broom::tidy(conf.int=TRUE)

# Margins
chol_margin = betamfx(highchol ~ walkable + sex_male + age_18plus + black + median_income, data = regress_df) %>% broom::tidy(conf.int=TRUE)
mental_margin = betamfx(mhlth ~ walkable + sex_male + age_18plus + black + median_income, data = regress_df) %>% broom::tidy(conf.int=TRUE)
lpa_margin = betamfx(lpa ~ walkable + sex_male + age_18plus + black + median_income, data = regress_df) %>% broom::tidy(conf.int=TRUE)
physical_margin = betamfx(phlth ~ walkable + sex_male + age_18plus + black + median_income, data = regress_df) %>% broom::tidy(conf.int=TRUE)
bloodpres_margin = betamfx(bphigh ~ walkable + sex_male + age_18plus + black + median_income, data = regress_df) %>% broom::tidy(conf.int=TRUE)

# Collate estimates into one data frame
modeldata = 
  chol %>% 
  filter(grepl('walk', term)) %>% 
  rbind(mental[2,]) %>% 
  rbind(lpa[2,]) %>% 
  rbind(physical[2,]) %>% 
  rbind(bloodpres[2,]) %>% 
  mutate(
    outcome = c("High Cholesterol", "Mental Health", "Physical Activity", "Physical Health", "High Blood Pressure"),
    OR = exp(estimate),
    lowlim = exp(conf.low),
    highlim = exp(conf.high)
  ) %>% 
  dplyr::select(outcome, OR, lowlim, highlim, p.value) %>% 
  arrange(desc(OR)) %>% 
  knitr::kable(digits = 3, col.names = c('Outcome', 'Coefficient', 'CI (lower)', 'CI (upper)', 'p-value'))

marginsdata = 
  chol_margin %>% 
  filter(grepl('walk', term)) %>% 
  rbind(mental_margin[1,]) %>% 
  rbind(lpa_margin[1,]) %>% 
  rbind(physical_margin[1,]) %>% 
  rbind(bloodpres_margin[1,]) %>% 
  mutate(
    outcome = c("High Cholesterol", "Mental Health", "Physical Activity", "Physical Health", "High Blood Pressure"),
    pct = estimate*100,
    pct.conf.low = conf.low*100,
    pct.conf.high = conf.high*100
  ) %>% 
  dplyr::select(outcome, pct, pct.conf.low, pct.conf.high, p.value) %>% 
  arrange(desc(pct)) %>% 
  knitr::kable(digits = 3, col.names = c('Outcome', '% change', 'CI (lower)', 'CI (upper)', 'p-value'))
```


#### Mental health

```{r, echo=FALSE}
mental %>% 
  dplyr::select(!c("component", "statistic", "std.error")) %>% 
  knitr::kable(digits = 3, col.names = c("Term", "Coefficient", "p-value", "CI (lower)", "CI (high)"))
```

#### Physical health

```{r, echo=FALSE}
physical %>% 
  dplyr::select(!c("component", "statistic", "std.error")) %>% 
  knitr::kable(digits = 3, col.names = c("Term", "Coefficient", "p-value", "CI (lower)", "CI (high)"))
```

#### Physical activity

```{r, echo=FALSE}
lpa %>% 
  dplyr::select(!c("component", "statistic", "std.error")) %>% 
  knitr::kable(digits = 3, col.names = c("Term", "Coefficient", "p-value", "CI (lower)", "CI (high)"))
```

#### High cholesterol

```{r, echo=FALSE}
chol %>% 
  dplyr::select(!c("component", "statistic", "std.error")) %>% 
  knitr::kable(digits = 3, col.names = c("Term", "Coefficient", "p-value", "CI (lower)", "CI (high)"))
```

#### High blood pressure

```{r, echo=FALSE}
bloodpres %>% 
  dplyr::select(!c("component", "statistic", "std.error")) %>% 
  knitr::kable(digits = 3, col.names = c("Term", "Coefficient", "p-value", "CI (lower)", "CI (high)"))
```

### Interpretation

By exponentiating the regression coefficient, we obtained the odds ratios for the association between health outcomes and walkability below. However, they are not at all straightforward to interpret, as they correspond to the odds of change in proportion of a selected health outcome, controlling for sociodemographic variables.

<details>
<summary>Show</summary>
```{r model, echo=FALSE}
modeldata
```
</details>
<br>

It is easier to convert the regression results into changes in marginal probabilities of the outcome variable:
```{r margins, echo=FALSE}
marginsdata
```

Here, we see that more walkable places have the following change in prevalence for a health outcome compared to less walkable places:

* 0.127% increase in poor mental health
* 0.206% decrease in no physical activity
* 0.627% decrease in poor physical health
* 3.226% decrease in poor cholesterol
* 4.367% decrease in poor blood pressure

Therefore, increase in walkability is associated with decreased prevalence of poor health outcomes, which is expected. However, an increase in walkability is associated with an increase for prevalence of poor mental health, but this could be due to unaccounted for confounding from density of space and urbanicity that comes with more walkable spaces.

### Forest Plot
Below, we create a forest plot to more visually display the difference in prevalence for our various health outcomes compared to less walkable places:
```{r forestplot-packages}
devtools::install_github("rdboyes/forester")

library(grid)
library(forester)
```

```{r forest-plot-df}
forestplot_df = 
  chol_margin %>% 
  filter(grepl('walk', term)) %>% 
  rbind(mental_margin[1,]) %>% 
  rbind(lpa_margin[1,]) %>% 
  rbind(physical_margin[1,]) %>% 
  rbind(bloodpres_margin[1,]) %>% 
  mutate(
    outcome = c("High Cholesterol", "Mental Health", "Physical Activity", "Physical Health", "High Blood Pressure"),
    pct = estimate*100,
    pct.conf.low = conf.low*100,
    pct.conf.high = conf.high*100
  ) %>% 
  dplyr::select(outcome, pct, pct.conf.low, pct.conf.high, p.value) %>% 
  arrange(desc(pct)) %>%
  janitor::clean_names()

forestplot_df
```


```{r plot}
forest_plot = forestplot_df %>%
  ggplot(aes(y = fct_rev(outcome))) + 
  theme_classic() +
  geom_point(aes(x = pct),
             size = 1) +
  geom_errorbarh(aes(y = outcome,
                     xmin = pct_conf_low,
                     xmax = pct_conf_high,
                     height = 0.1)) +
  geom_vline(xintercept = 0,
             linetype = "dashed") +
  labs(title = "Forest Plot",
       subtitle = "Estimated % Change in Health Outcomes",
       x = "% Change in Prevalence (with 95% CI)",
       y = "Health Outcomes") +
  theme(plot.title = element_text(size = 13, hjust = 0.5),
        plot.subtitle = element_text(size = 11, hjust = 0.5),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.position = "none") +
  xlim(-5, 2) +
  annotate("text", x = -1, y = 5, label = "Decrease", size = 3) +
  annotate("text", x = 1, y = 5, label = "Increase", size = 3) +
  coord_cartesian(xlim = c(-4.5, 1.5))

forest_plot
```



