---
title: "Data Cleaning"
output: github_document
---

<style>
div.blue { background-color:aliceblue; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

##### This is a description of our data cleaning process.

##### **Due to the size of the original data files, please download the datasets according to the links below and save the .csv files locally in its original name. And save all .csv file in a new folder called "Data" in order to run the code.**

</div>

<br>

##### Walkability

* Data: [Environmental Protection Agency's information on Walkability Index, 2021](http://edg.epa.gov/metadata/catalog/search/resource/details.page?uuid=%7B251AFDD9-23A7-4068-9B27-A3048A7E6012%7D)
* Cleaning steps:
  + Added leading zeros for all Federal Information Processing Standard (FIPS) codes (numbers which uniquely identify geographic areas)
  + Combined state, county and tract FIPS into an 11-digit code
  + Selected relevant variables
* Variables of interest:
  + FIPS code
  + National Walkability Index 

<details>
  <summary>Click to show code</summary>
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)

```


```{r}
#Added leading zeros for all Federal Information Processing Standard (FIPS) codes
#Combined state, county and tract FIPS into an 11-digit code 
walkability_raw = read_csv("./data/EPA_SmartLocationDatabase_V3_Jan_2021_Final.csv")

walkability_new = walkability_raw %>% 
  janitor::clean_names() %>% 
  mutate(statefp = str_pad(statefp, width = 2, pad = "0"),
         countyfp = str_pad(countyfp, width = 3, pad = "0"),
         tractce = str_pad(tractce, width = 6, pad = "0"),
         fips = str_c(statefp, countyfp, tractce, blkgrpce, sep = "")) %>% 
  select(fips, everything(), -objectid)

#Selected relevant variables
walkability_clean = walkability_new %>% 
  select(fips, statefp, countyfp, tractce, blkgrpce, nat_walk_ind, csa_name) %>% 
  mutate(location_name = str_c(statefp, countyfp, tractce, sep = ""))
  
```
</details>
<br>

##### Demographics
* Data: [US Census American Community Survey, 2020](https://data.census.gov/table?tid=ACSDP1Y2021.DP05)
* Cleaning steps:
  + removed second row with detailed column names
  + removed columns for annotations and margins of error
  + created a census tract column derived from the geo_id column (removed the "1400000US" at the beginning of each entry) to match with other data sets
* Variables of interest:
  + Age
  + Sex
  + Race/ethnicity
  + Median household income

<details>
  <summary>Click to show code</summary>
  
Demographic Data - ACS 2020
Age, Gender, and Race
```{r}
demo_raw = read_csv("./data/ACSDP5Y2020.DP05-Data.csv")

#removed second row with detailed column names
#removed columns for annotations and margins of error
#created a census tract column derived from the geo_id column (removed the "1400000US" at the beginning of each entry) to match with other data sets
demo_clean = demo_raw %>% 
  slice(-1) %>% 
  janitor::clean_names() %>%
  select(-ends_with(c("ea", "m", "ma"))) %>% 
  mutate(geo_clean = geo_id, 
         geo_clean = str_replace(geo_clean, "^1400000US", "")) %>% 
  select(geo_clean, dp05_0002pe, dp05_0003pe, dp05_0021pe, dp05_0037pe, dp05_0038pe, dp05_0039pe, dp05_0044pe, dp05_0058pe, dp05_0071pe)

```

Median household income 
```{r}
income_raw = read_csv("./data/ACSST5Y2020.S1903-Data.csv")

income_clean = income_raw %>% 
  janitor::clean_names() %>%
  select(geo_id, name, s1903_c03_001e) %>% 
  slice(-1) %>% 
  mutate(geo_clean = geo_id, 
         geo_clean = str_replace(geo_clean, "^1400000US", "")) %>% 
  select(geo_clean, everything())

```

</details>
<br>

##### Health outcomes
* Data: [Behavioral Risk Factor Surveillance System (BRFSS), 2020](https://chronicdata.cdc.gov/500-Cities-Places/PLACES-Local-Data-for-Better-Health-Census-Tract-D/cwsq-ngmh/data)
* Cleaning steps:
  + Removed footnote variables
  + Filtered for relevant health outcomes
* Variables of interest:
  + High cholesterol
  + Physical health
  + Blood pressure
  + Physical inactivity
  + Mental health
  
<details>
  <summary>Click to show code</summary>

Health Outcome Data - BRFSS
```{r}
# raw data with all health outcomes
health_df = read_csv("./data/PLACES__Local_Data_for_Better_Health__Census_Tract_Data_2022_release.csv")

health_new = health_df %>% 
  janitor::clean_names() %>%
  mutate(location_name = str_pad(location_name, width = 11, pad = "0"),
         geolocation = str_replace(geolocation, "^POINT ", ""),
         geolocation = str_replace(geolocation, "^\\(", ""),
         geolocation = str_replace(geolocation, "\\)$", "")) %>% 
  separate(geolocation, into = c("lat", "long"), sep = " ")
  
```

Further cleaning
```{r}

#For health outcome, we are focusing on 5 outcomes that are likely to be related to walkability 
health_clean = health_new %>% 
  filter(measure_id == c("BPHIGH", "PHLTH", "LPA", "MHLTH", "HIGHCHOL")) %>% 
  select(location_name, lat, long, measure, data_value, data_value_type, data_value_unit, measure_id)
```

</details>
<br>

##### Cleaned Dataset for Analysis Use
After cleaning all three data sources, we merged the dataset by geographical locations. 

<details>
  <summary>Click to show code</summary>
  
Merge
```{r}
# merging walkability and health outcomes
merge1 = full_join(x = walkability_clean, y = health_clean, by = "location_name")

#mergeing demographic into one df - age&race + median household income
merge2 = full_join(demo_clean, income_clean, by = "geo_clean")
  
merge2 = merge2 %>% 
  mutate(location_name = geo_clean) 

# merge merge1 and merge2
merge_final = full_join(x = merge1, y = merge2, by = "location_name")

merge_final = merge_final %>% 
  rename(sex_male = dp05_0002pe,
         sex_female = dp05_0003pe,
         age_18plus = dp05_0021pe,
         race_white = dp05_0037pe,
         race_black = dp05_0038pe,
         race_aian = dp05_0039pe,
         race_asian = dp05_0044pe,
         race_2_plus = dp05_0058pe,
         race_hispanic = dp05_0071pe,
         median_income = s1903_c03_001e) %>% 
  pivot_longer(names_to = "race",
               values_to = "percent_race",
               cols = starts_with("race"),
               names_prefix = "race_") %>% 
  select(-data_value_type, -data_value_unit)

write_csv(merge_final, "./data/merge_final.csv")

```

</details>


